---
title: "Project Proposal Testing"
author: "Asmit Chakraborty"
output:
    pdf_document: default
fontsize: 11pt
---

```{r}
beats <- read.csv("tx_houston_shapefiles/Houston_Police_Beats.csv", header=TRUE)
districts <- read.csv("tx_houston_shapefiles/Houston_Police_Districts.csv")
divisions <- read.csv("tx_houston_shapefiles/Houston_Police_Division.csv")
df <- read.csv("tx_houston_2023_01_26.csv")
```

```{r}
summary(df)

table(df$subject_race)
table(df$subject_sex)
table(df$type) # All vehicular
table(df$outcome) # All citation
table(df$district) # Could be interesting
table(df$citation_issued) # All TRUE
summary(df$speed) # about 30% of obs have speed tracked, what is speed vs posted speed?
```

Houston seems kind of boring. All stops received a cittion?

Potentially interesting problems
- clustering "high police stop risk" areas?
- Anything on differences between speeds in houston vs austin?

Newer (but have not looked closely at data yet for these)
- Clearly there are racial disparities in police stops. Are these more influenced
by who is stopped (age, race, sex, etc) or where the stop happened (district, etc)
  - Fit some sort of mixed effects model?

- Do stops/searches change based on day / night?
  - Idea is that, at night, police have a harder time profiling people

- For similar violations, are any demographic characteristics associated with a higher
likelihood of getting a citation or going to jail?

```{r}
df2 <- read.csv("~/Downloads/California.csv")
View(df2)
```


FIPS is the tract code

E_HU is the like estimated # of houses in that tract

so theres like house-level data which should lead us to be able to combine
tracts inside each of the like precincts or whatever those policing zones were
in that map i sent earlier

I think

Then we can take an average across all the tracts in each policing zone 

and if the SOPP data row {i} sits in precinct {j}, then the SVI column for row {i}
should be like the average of all the tract SVIs in precinct {j}

also like thats my ideal guess of how this would work out without having done any of it yet


do i keep sharing screen....